Human-level control through deep reinforcement learning

### 2018-01-07 article/article.tex
Get further into the article

\usepackage[utf8]{inputenc}
\usepackage[sort, colon]{natbib}
\usepackage{alifexi}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{commath}

@@

\title{Human-level control through deep reinforcement learning}
\author{C\'{e}dric Simar, Antoine Passemiers, Stanislas Gueniffey \and Robin Petit \\
\mbox{}\\
Universit\'{e} Libre de Bruxelles \\
\{Cedric.Simar, apassemi, sgueniff, robpetit\}@ulb.ac.be}

@@

The huge breakout effect of the DQN was that it has shown to be a first step towards general artificial intelligence~\citep{togelius2015ai}, meaning an
algorithm that is not only good at learning one particular task but several of them. Even though the DQN needs to be trained for about a million steps
on each Atari game, it gets good at most of them and better than human at 22 of the 49 experimented~\citep{mnih2015human}.

Also we compared DQN, as well as other state-of-the-art deep reinforcement learning architectures, to simpler reinforcement learning techniques (including
random action selection) using a normalized performance metric. The latter is formalized as below:
\begin{equation}
    P = 100 \times (\text{score} - RAS) / (HS - RAS),
\end{equation}
where P designates the normalized performance, RAS the random action selection score, and HS the human score obtained in average by our
Space Invaders professional gamer \textbf{Stanislas Gueniffey}. By normalizing in this manner, the random action selection score RAS results in
a normalized performance of 0\% and the human score results in a normalized performance of 100\%.

@@

@@

The interest of the Hubert loss is to bound the gradient of the loss between $-1$ and $+1$, and has (\todo {probably}) been shown to increase
stability.~\todo{check if this has a proper reference or drop it}

@@

These are then reshaped into a column vector (of size $7 \times 7 \times 64 = 3136$) and fed into the first fully-connected layer with 512 neurons,
which is then connected to the last layer (again fully-connected) which corresponds to the output with the same size as the action space (6 in the
case of Space Invaders).

@@

The last layer of the DDQN outputs Q-value estimates by evaluating the following formula:
\begin{multline}
    Q(s, a; \theta, \alpha, \beta) = V(s; \theta, \beta) + \\ ( A(s, a; \theta, \alpha) - \frac{1}{|\mathcal{A}|} \Sigma_{a'} A(s, a'; \theta, \alpha) )
\end{multline}
where $V$ is the output of the state-value stream, $A$ is the output of the advantage-per-action stream, $s$ is the input state (image), 
$\alpha$ are the parameters of the state-value stream's fully-connected layer, $\beta$ are the parameters of the state-value stream's
fully-connected layer, $\theta$ are the parameters of the rest of the network, and $\mathcal{A}$ is the discrete action space.

@@

Every 4 steps, at time $t$, a minibatch of 32 past experiences $\left\{(s_{t_j}, a_{t_j}, r_{t_j}, s_{t_j+1})\right\}_{j=1}^{32}$ is selected uniformly from the
memory in order to perform a Stochastic Gradient Descent (SGD), thus the average of the gradients of the loss function on each experience is used to update
the weights. As mentioned above, the loss function is a Huber loss function and is defined as:
\begin{equation}
	H_1\left(y_j - Q(s_{t_j}, a_{t_j}; \theta)\right),
\end{equation}
with $y_j$ set either to $r_j$ if $s_{t_j}$ is a final state (i.e. either the loss of a game or a forced end of episode) or to
$r_j + \gamma\max_{a' \in A(s_{t_j})}\hat Q(s_{j+1}, a'; \theta^-)$ otherwise; therefore using the target action-value network.

@@

\begin{figure*}
  \includegraphics[width=\textwidth]{figures/dqn_uniform}
  \caption{(Left) Total score per episode during a learning phase of 1.000.000 steps, smoothed using a moving average with a window of size 2000.
     (Right) Estimation of Q, averaged across actions, for each batch during a learning phase of 1.000.000 steps and smoothed using a moving average
      with a window of size 2000.}
\end{figure*}

Figure~\ref{fig:t-SNE}\todo{Link the right figure} shows a 2-dimensional representation of the last fully connected hidden layer, i.e. a 512-dimensional
vector. The method of \textit{projection} that was used is t-SNE~\citep{maaten2008visualizing}, which is part of the manifold learning algorithms, also
called \textit{Nonlinear Dimensionality Reduction}~\citep{lee2007nonlinear}.

@@

  \todo{Magnifique plot T-SNE avec des couleurs et des gifs et des trucs qui font beep et qui font boop}~\citep{wattenberg2016how}

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage[sort, colon]{natbib}
\usepackage{alifexi}
\usepackage{float}
\usepackage{subcaption}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{commath}

@@

\title{Human-level control through deep reinforcement learning}
\author{C\'{e}dric Simar, Antoine Passemiers, Stanislas Gueniffey \and Robin Petit \\
\mbox{}\\
Universit\'{e} Libre de Bruxelles \\
\{csimar, apassemi, sgueniff, robpetit\}@ulb.ac.be}

@@

The huge breakout effect of the DQN was that it has shown to be a first step towards general artificial intelligence~\citep{togelius2015ai}, meaning an
algorithm that is not only good at learning one particular task but several of them. Even though the DQN needs to be trained for tens of millions steps
on each Atari game, it gets good at most of them and better than human (considered $75\%$ of human score or more) at 22 of the 49 experimented~\citep{mnih2015human}.

@@

Also we compared DQN, as well as other state-of-the-art deep reinforcement learning architectures, to simpler reinforcement learning techniques (including
random action selection) using a normalized performance metric. The latter is formalized as below:
\begin{equation}
    P = 100 \times (\text{score} - RAS) / (HS - RAS),
\end{equation}
where P designates the normalized performance, RAS the random action selection score, and HS the human score obtained in average by Google's professional
game tester after a training period of 2 hours.  By normalizing in this manner, the random action selection score RAS results in a normalized performance
of 0\% and the human score results in a normalized performance of 100\%.

@@

The interest of the Hubert loss is to bound the gradient of the loss between $-1$ and $+1$, and is claimed to increase stability~\citep{mnih2015human},
yet no further investigation has been performed.

@@

These are then reshaped into a column vector (of size $7 \times 7 \times 64 = 3136$) and fed into the first fully-connected layer with 512 neurons,
which is then connected to the last layer (again fully-connected) which corresponds to the output with the same size as the action space (4 in the
case of Breakout: \textit{NOOP}, \textit{FIRE}, \textit{LEFT} and \textit{RIGHT}).

@@

The last layer of the DDQN outputs Q-value estimates by evaluating the following formula:
\begin{multline}
    Q(s, a; \theta, \alpha, \beta) = V(s; \theta, \beta) + \\ \left(A(s, a; \theta, \alpha) - \frac{1}{|\mathcal{A}|} \sum_{a'} A(s, a'; \theta, \alpha)\right)
\end{multline}
where $V$ is the output of the state-value stream, $A$ is the output of the advantage-per-action stream, $s$ is the input state (image),
$\alpha$ are the parameters of the state-value stream's fully-connected layer, $\beta$ are the parameters of the state-value stream's
fully-connected layer, $\theta$ are the parameters of the rest of the network, and $\mathcal{A}$ is the discrete action space.

@@

Every 4 steps, at time $t$, a minibatch of 32 past experiences $\left\{(s_{t_j}, a_{t_j}, r_{t_j}, s_{t_j+1})\right\}_{j=1}^{32}$ is selected uniformly from the
memory in order to perform a Stochastic Gradient Descent (SGD), thus the average of the gradients of the loss function on each experience is used to update
the weights. As mentioned above, the loss function is a Huber loss function and is defined as:
\begin{equation}
	H_1\left(y_j - Q(s_{t_j}, a_{t_j}; \theta)\right),
\end{equation}
with $y_j$ set either to $r_{t_j}$ if $s_{t_j}$ is a final state (i.e. either the loss of a game or a forced end of episode) or to
$r_{t_j} + \gamma\max_{a' \in A(s_{t_j})}\hat Q(s_{t_j+1}, a'; \theta^-)$ otherwise; therefore using the target action-value network.

@@

\begin{figure*}[!t]
	\vspace{-1cm}
	\begin{subfigure}{.47\textwidth}
		\includegraphics[width=\textwidth]{figures/dqn_uniform_e_scores}
		\subcaption{Total score per episode during a learning phase of 16,000,000 steps, smoothed using a moving average with a window of size 1000.}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.47\textwidth}
		\includegraphics[width=\textwidth]{figures/dqn_uniform_q_values}
		\subcaption{Estimation of $Q$, averaged across actions, for each batch during a learning phase of 16,000,000 steps and smoothed using a moving
			average with a window of size 5000.}
	\end{subfigure}
	\caption{Episode score and $Q$ values estimation of the DQN after 16,000,000 steps of training on Breakout.\label{fig:dqn uniform}}
\end{figure*}

Figure~\ref{fig:dqn uniform} shows the results of the DQN after 16,000,000 steps of training. One can observe that the first 15,000 batches didn't help
increasing the $Q$ estimates: the internal representation was being built up. Also, the score per episode is very low at first since the agent plays
randomly, a million steps were required for $\epsilon$ to linearly decrease from $1$ to $0.1$.

After this phase, the estimates increase and so does the score per episode, until they both reached stability around 11,000,000 steps.

\begin{table}[!h]
	\centering
	% std of Random is 1.27 & std of DQN is 22.81
	\begin{tabular}{l|l|l|l}
		Model & Random & Human & DQN \\ \hline \hline
		Score & $1.31$ & $31.8$ & $39.88$ \\ \hline
		P &  $0\%$ & $100\%$ & $126.83\%$
	\end{tabular}
	\caption{Score and Performance on Breakout (averaged on 100 simulations with $\varepsilon=5\%$ and with one action decision per frame). The DQN has been
	trained during 16,000,000 steps.
	\todo{Ajouter scores Prioritized + DDQN et le temps d'apprentissage}\label{tab:scores breakout}}
\end{table}

Table~\ref{tab:scores breakout} shows that the DQN performed better than human after 16,000,000 training steps, with a performance of $126.83\%$.

Figure~\ref{fig:t-SNE DQN uniform} shows a 2-dimensional representation of the last fully connected hidden layer, i.e. a 512-dimensional
vector. The method of \textit{projection} that was used is t-SNE~\citep{maaten2008visualizing}, which is part of the manifold learning algorithms, also
called \textit{Nonlinear Dimensionality Reduction}~\citep{lee2007nonlinear}.

@@

  \todo{Magnifique plot T-SNE avec des couleurs et des gifs et des trucs qui font beep et qui font boop. Robin est dessus.}~\citep{wattenberg2016how}

\begin{figure}
	\includegraphics[width=.5\textwidth]{figures/TSNE_dqn_uniform}
	\caption{\todo{Replace with correct plot}.\label{fig:t-SNE DQN uniform}}
\end{figure}

Figure~\ref{fig:dqn uniform scores dist} shows the distribution of scores when letting the DQN play for 100 games. We can see that despite having an
average score of \todo{XXX}, the DQN outputs score ranging from 7 to 219. This means that some states haven't been encountered enough for the
DQN to know significantly which one is better, and can also be slightly related to the $\epsilon$-greedy policy, yet $\epsilon$ is quite small,
and more importantly, the agent selects the action at each frame. This highest score of 219 is probably due to a non-anticipated breaking through
the brick layer, leading to the ball destroying the upper layer.

\begin{figure}
	\includegraphics[width=.5\textwidth]{figures/dqn_uniform_score_distribution}
	\caption{Distribution of the scores performed by the DQN on 100 plays.\label{fig:dqn uniform scores dist}}
\end{figure}

\begin{figure*}{!b}
	\begin{subfigure}{.47\textwidth}
		\includegraphics[width=\textwidth]{figures/dqn_uniform_e_scores}
		\subcaption{Total score per episode of the DDQN during its learning phase, smoothed using a moving average of size \todo{XXX}.}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.47\textwidth}
		\includegraphics[width=\textwidth]{figures/dqn_uniform_q_values}
		\caption{Estimation of $Q$ by the DDQN, averaged across actions, for each batch during its learning phase and smoothed using a moving average
		with a window of size \todo{XXX}.}
	\end{subfigure}
	\caption{\todo{Ajouter les plots DDQN et remplacer les paths des includegraphics}.\label{fig:ddqn uniform}}
\end{figure*}

@@

\section{Conclusion}

	\todo{Ajouter ordre de performance des 3 tests}
